Understanding the nanoscale structure of hexagonal phase lyotropic liquid crystal membranes

### 2017-12-05 Ben_Manuscripts/structure_paper/Outline.tex
Minor comments?
comments on latest outline.

  While RO and NF have seen many advances in the past few decades, they 
  are far from perfect separation technologies.
  \begin{itemize}
    \item \textit{RO membranes}
    \item Inconsistent performance : Current state-of-the-art RO membranes are unstructured with
    tortuous and polydisperse diffusion pathways which leads to 
    inconsistent performance \cite{song_nano_2011}
    \item High energy requirements : Necessarily high feed pressures 
    drive up energy requirements which strains developing regions and
    contributes strongly to CO\textsubscript{2} emissions. \cite{mcginnis_global_2008}
    \item Separation based on differences in solubility and diffusivity:
%MRS2: couldn't you argue that the fact that they diffuse and dissolve at different rates means that it makes it possible to design targeted separations?
%BJC: right, I say something like that in the sentence following this one. It would be more challenging to design a targeted separation 
% especially if the feed is complex
    Moreover, designing RO membranes to achieve targeted separations of 
    specific solutes in a complex feed solution is nearly impossible because various solutes dissolve
    into and diffuse through the polymer matrix at different rates. \cite{wijmans_solution-diffusion_1995}
    \item With optimization, one can exploit these differences to create a functional
    selective barrier.
    \item \textit{NF membranes}
    \item NF was introduced as an intermediate between RO and ultrafiltration,
    having the ability to separate organic matter and salts on the order of 
    one nanometer in size.
%MRS2: is ``well-defined'' the right word if there is a ``pore size distribution''?
%BJC: By well-defined, I'm thinking of cylinders versus polymer matrix. One is definitely more
% defined than the other. What is the right word for that? unambiguous? explicit?
    \item Larger and well-defined pores drive down energy requirements while
    still affording separation of solutes as small as ions to some degree \cite{van_der_bruggen_review_2003}
    \item NF is often used as a precursor to reverse osmosis
    \item Unfortunately, NF membranes, like RO, possess a pore size 
    distribution which limits their ability to perform precise separations \cite{bowen_modelling_2002}
  \end{itemize}

@@

  A molecular level understanding of LLC membrane structure, enabled by molecular
  dynamics simulations, will provide guidelines to reduce the large chemical space
  available to design monomers for creation of separation-specific membranes.
  \begin{itemize}
    \item Over the past 20 years, H\textsubscript{II} phase LLC membrane studies have been limited primarily
    to Na-GA3C11 with some characterization done after minor structural modifications
    \item Resel et al. varied the length of the monomer tails and the counterion used and 
    observed its affect on pore spacing \cite{resel_structural_2000}.
    \item Rejection studies show that this membrane can not perform separations of
    solutes less than 1.2 nm in diameter because the pores are too large \cite{zhou_supported_2005}.
    \item  We do not yet understand how to controllably reduce the effective pore size or 
    how to tune the chemical environment in the nanopores for effective water
    desalination or small organic molecule separations.
    \item It will be challenging to efficiently narrow down the large design space in 
    a laboratory setting without a robust model.
    \item The only source of predictive modeling for LLC systems have been macroscopic models 
    which likely do not adequately describe transport at these length scales. % BJC: Reference. I think it is just w.r.t. bicontinuous cubic
%MRS2: not quite sure of the phrasing here on next two points. Something more a long the lines of the choice of head group 
%may allow tuning rejection of solutes based on size?
%BJC: okay tried again
    \item The choice of head group may allow us to tune pore size for size exclusion driven separations
    \item Choice of counterion may influence the establishment of a Donnan potential
    affecting the degree to which the membrane can exclude charged species.
    \item A good molecular model should incorporate a detailed picture of the nanoscopic pore 
    structure which will be crucial to understanding the role of monomer structure in 
    membrane design.
    \item Molecular dynamics simulations will provide the required level of detail
  \end{itemize}

  %BJC: Figure here showing previous level of detail compared to atomistic level of detail
  %MRS2: provenance of the atomistic figure could be explained better in the caption. It also looks disordered. Will eventually need a higher resolution figure for the previous understanding pore.
  %BJC: Okay. The picture here is of the "ordered pore" phase. It looks disorderd from this angle. I could use a different slice of the membrane to see if it looks any better
  \begin{figure}
  \centering
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[width=\textwidth]{nanopore_undetailed.jpg}
		\caption{}~\label{fig:undetailed_pore}
	\end{subfigure}
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[width=\textwidth]{detailed_pore.png}
		\caption{}~\label{fig:detailed_pore}
	\end{subfigure}
  \caption{(a) Our previous understanding of the pore structure allows us to speculate
	   about separation behavior. (b) A detailed molecular model will allow us to
	   directly observe solute transport. Here, four stacked layers of 5 monomers
           are pictured atomistically. The hydrophilic region is in red and the 
	   hydrophoic region is colored blue.}~\label{fig:detail}
  \end{figure}

@@

  Despite having structural data, there is still information which
  experiment cannot definitively answer. There are several key
  questions that we intend to answer which will be laid out and numbered
  in subsequent paragraphs.

@@

%MRS: break this down into more hierarchy - there are 11 separate points.  Can they be grouped at all for clearer understanding? elucidate what the aspects are we need a clearer picture. 
%BJC: I moved the logic for choice of monomer to system to model to a separate paragraph (above this). I then moved the other three things I talk about into their own paragraphs to keep the ideas separate (following paragraphs).

%MRS: since the questions we answer in the paper have a lot to do with the structure of the membranes, somewhere relatively early in the paper we need to lay out the facts that 1) these membranes could be really good (I think you do establish this) 2) we really would need to know the structure better to really do useful things with these, and 3) MD is a very helpful way to gain more information about the structure, because of the limited range of experiments we can do with these systems. That is the main rationale, and it should be made clear fairly early in the introduction.  Very important to make your story absolutely as clear as possible.
%BJC: I moved and edited a paragraph that was originally later on to be 2 above this one. I think it does a better job addressing points 2 and 3. And now it addresses those points earlier in the intro (as early as it could go where I think it still makes sense in context)

@@

  Monomers in the Col\textsubscript{h} system are theorized to be partitioned
  into stacked layers which form columnar pores. There has been no definitive
  answer in literature regarding the number of monomers in each layer. We want to
  know 
\begin{enumerate}
\item If layers do exist, how many monomers constitute a single layer? \label{point:monomernum}
  \begin{itemize}
    \item A simple molecular simulation study of a similar molecule suggested
    that there are 4 monomers in each layer. Their estimation is based on a
    simulated system containing only 16 total monomers which likely does not sufficiently
    model the chemical environment present in the real system.~\cite{zhu_methacrylated_2006}. 
    \item A separate calculation based on the volume of the liquid crystal monomers proposes
    that there are seven monomers in each layer~\cite{resel_structural_2000}. 
    \item A molecular model orders of magnitude larger than any other reported atomistic 
    liquid crystal membrane simulations has the best chance of directly answering this question.
    \item We can directly change the layer composition and note its effect on membrane structure.
  \end{itemize}

@@

  \item Is it necessary to include any water in order to appropriately model the 
  Col\textsubscript{h} phase? \label{point:water}
  \begin{itemize}
	\item While the Col\textsubscript{h} phase is described as dry, it is 
	possible that small amounts of ambient water may be leached into the system.
	\item The hydrogen bonding network formed by the water may play a role 
        in structuring the pore.
	\item We can use simulated X-ray diffraction patterns to see if there is
        any meaningful structural difference between a "dry" and "wet" system
  \end{itemize}
  \end{enumerate}

@@

  We chose the layer spacing for the initial configuration based on R-$\pi$ 
  and then allowed system to readjust during equilibration.
  \begin{itemize}  
    \item Each monomer was rotated so the plane of the aromatic head groups would
    be coplanar with the xy plane.
    \item We explore two different initial layer spacings
    \item The first is exactly equal to R-$\pi$ with layers placed so aromatic
    rings are stacked 3.7 \angstrom~ apart in the z-direction.
    \item A second system is explored with an initial layer spacing of 5 \angstrom
    \item A third system with an initial layer spacing of 10 \angstrom was briefly
    explored and the reader can learn more about it in the Supplemental Information.
% BJC: Move all this lower down to be with the rest of the equilibrium calculations
%    \item To quantify the degree of layering in our system, we calculated
%    a spatial correlation function, $g(z)$, measured along the z-axis 
%    (perpendicular to the membrane plane)
%    \item To calculate $g(z)$, we binned the z component distances between 
%    the center of mass of each benzene ring and all others of the same pore 
%    over 50 ns of equilibrated trajectory and then normalized by the average
%    number density.
%    \item We compare the degree of layering between systems based on the
%    amplitude of the first peak in $g(z)$. We halve the difference between the
%    maximum of the first peak and the following minimum. We compare the difference
%    to the mean to get a percentage deviation from the average number density.
%    \item To extract the average distance between layers we applied a discrete 
%    fourier transform to $g(z)$ and extracted the highest intensity frequency 
  \end{itemize}

@@

  We developed equilibration schemes to create dry and wet configurations. Both schemes 
  start with an initial configuration generated according to the previous guidelines.
  \begin{itemize}
      \item Dry equilibration scheme:
      \begin{itemize}
        %MRS2: what initial force constant, and does it matter (that can be in supplemental) 
	%BJC: I've never explored whether there is any dependence on initial force constant. It probably doesn't matter unless it starts really low
          \item Restraints fix monomer head groups in the sandwiched or parallel-displaced
          configurations while allowing monomer tails to settle.
          \item Doing so also mitigates system dependence on initial monomer configuration.
          \item The restrained portion of the equilibration scheme is run in the NVT ensemble.
          \item Every 50 ps, we reduce the force constants by the square root of its
          previous value, starting from 1e6 KJ mol$^-1$ nm$^-2$.
          \item Once the force constant is below 10 KJ mol$^-1$ nm$^-2$, the restraints are
          released linearly until there is no more restraining potential.
          \item The resulting unrestrained structure is allowed to equilibrate for 5 ns 
	  in the NPT ensemble with pressure controlled by the berendsen barostat.
	  \item Long, NPT equilibration simulations are run for at least 400 ns using the 
          Parrinello-Rahman barostat with a time consant of 10 ps.
      \end{itemize}
      \item Wet equilibration scheme: 
      \begin{itemize}
	  % BJC: the following three bullet points ended up not being the case since I used significantly less water
          %\item We obtained an estimate of equilibrium pore water content by solvating
          % an initial configuration with water baths above and below the membrane. 
	  %\item We ran the solvated system according to Scheme A but with a 1000 ns equilibration
          %\item Using that number we added water to the pores of an initial 
	  %configuration and equilibrated according to scheme B. 
	  \item In order to create a wet system, we solvated an initial configuration with water using gmx solvate
          \item All water molecules placed outside the pore region are removed 
	  \item Waters inside the pore region are randomly removed until the desired concentration
	  of water in the pores is reached. 
	  \item The remainder of the equilibration is the same as the dry system
      \end{itemize}
      \item In all cases, the v-rescale thermostat was used with tau-t = 0.1 % BJC: supplemental
  \end{itemize}

@@

  To calculate the equilibrated pore spacing, we measured the distance between pore
  centers.
  \begin{itemize}
	\item Pore centers were located by averaging the coordinates of sodium ions in their
	respective pores.
	\item Statistics were generated using the bootstrapping technique (See Supplemental Information)
	% Supplemental vvvv
        \begin{itemize}
        	%MRS2: make clear sampling what?
	        \item We are interested in 5 pore-to-pore distances which should all be equal in
	        a perfect hexagonal array, however only 4 distances are independent % can visualize this in supplemental material. Will make a lot of things more clear below.
	        \item Each pore spacing has its own trajectory of spacing vs. time. Using data collected after
	        the system is equilibrated, we calculate how long it takes for the data in each of the 5
	        trajetories to become uncorrelated using pymbar.timeseries.integratedAutocorrelationTime() % reference
	        \item We break the full trajectories down into sub-trajectories based on the
	        maximum autocorrelation time of those found in the previous step.
	        \item For each bootstrap trial, we recreate an equilibrium trajectory by randomly
	        sampling pore spacings from the sub-trajectories
	        \item We get an average value for each pore spacing by finding the mean of the bootstrapped data
	        \item We calculate the overall average as the mean of all bootstrapped pore spacings
	        \item The uncertainty for each pore spacing is calculated as $\dfrac{<x> - x}{4}$ where $<x>$ is the
	        average spacing from the bootstrap trial and $x$ is the average value of one of the pore spacings.
	        \item We report the mean of these uncertainties
	        % MRS: need to deal with the fact only 4 are independent in the bootstrapping?  How did we do that?
	        % BJC: Updated
        \end{itemize}
        % Supplemental ^^^^^
  \end{itemize}

@@

  \begin{figure}
  \centering
  	\begin{subfigure}{\linewidth}
	\centering
		\begin{subfigure}{0.45\textwidth}
        		\centering
        		\includegraphics[width=\linewidth]{offset_tail_packing.png}
        		\caption{}~\label{fig:offset_tails}
		\end{subfigure}
		\begin{subfigure}{0.45\textwidth}
		\centering
	        	\includegraphics[width=\linewidth]{offset_angle_v_I.png}
		        \caption{}~\label{fig:offset_integration}
		\end{subfigure}
	\end{subfigure}
	\begin{subfigure}{\linewidth}
	\centering
		\begin{subfigure}{0.45\textwidth}
	        \centering
		        \includegraphics[width=\linewidth]{angles_traj_layered.png}
		        \caption{}~\label{fig:rz_layered}
		\end{subfigure}
		\begin{subfigure}{0.45\textwidth}
        	\centering
		        \includegraphics[width=\linewidth]{layered_angle_v_I.png}
		        \caption{}~\label{fig:layered_integration}
		\end{subfigure}
	\end{subfigure}
  \caption{The distribution of angles w.r.t. the xy plane between alkane chain tail centroids and nearest
  neighbor centroids for equilibrated parallel displaced (a) and sandwiched (c) configurations. The
  same peaks are visible when the 2D simulated diffraction data is radially integrated in the R-alkanes region,
  (b) and (d) respectively.}~\label{fig:tail_packing}
%MRS2: not clear what the difference is between right and left side figures.
%BJC: That's good! One is angles measured between tail centroids and the other is radial integration of the simulated xrd patterns. I'll try make that more clear in the caption.
  \end{figure}

  The $z$-direction correlation functions show that layers in our model prefer
  to stack further apart than the 3.7 \angstrom~ suggested by experiment. We attempted equilibration
  with layers stacked 5.0 nm apart and to our surprise, we observed
  long-term stability of a qualitatively different configuration suggesting
  that we have found more than one metastable free energy basin, corroborating \ref{point:metastable}.
  %that the initial distance between layers can influence the approach towards 
  %equilibrium.
  \begin{itemize}
        \item Equilibrated systems built according to the 3.7 \angstrom~layer spacing
        implied by R-$\pi$ are characterized by a defined, cylindrical and open pore
        structure.
        \item We will refer to this large set of configurations, with an open pore, as
        the ordered pore basin open pore (Figure~\ref{fig:OPbasin}).
        \item Simulations of systems built with layers stacked 5
        \angstrom~ apart results in a pore structure characterized by high radial
        disorder, while still maintaining partitioning between hydrophobic and
        hydrophilic regions.
        \item This will be called the disordered pore basin (Figure~\ref{fig:DPbasin}).
        \item This LLC membrane may exist in at least two metastable states.
        \item The distinct difference in pore structure exhibited by each phase will
        likely lead to different transport mechanisms.
        \item Because the pore structure will vary between each Basin, understanding
        which phase exists experimentally is necessary in order to ensure we are
        studying the system which actually dominates.
  \end{itemize}

  \begin{figure}[!ht]
        \centering
        \begin{subfigure}[b]{0.475\textwidth}
                \centering
                \includegraphics[width=\textwidth]{280K_tramp_close.png}
                \caption{}\label{fig:OPbasin}
        \end{subfigure}
        \begin{subfigure}[b]{0.475\textwidth}
                \centering
                \includegraphics[width=\textwidth]{340K_tramp_close_full.png}
                \caption{}\label{fig:DPbasin}
        \end{subfigure}
        \caption{From a qualitative standpoint, the ordered pore basin (a) is characterized by a
        hollow cylindrical pore while pores in the disorder pore basin (b) are characterized by a
        higher degree of radial disorder}\label{fig:basins}
  \end{figure}

@@

  % BJC: I'm not sure what to make of this figure right now. The parallel-displaced configuration shares
  % more with the disordered pore phase
  \begin{figure}
  \centering
  \begin{subfigure}{0.55\textwidth}
        \includegraphics[width=1\linewidth]{offset_density.png}
        \caption{}
        \label{fig:offset_density}
  \end{subfigure}
  \begin{subfigure}{0.55\textwidth}
        \includegraphics[width=1\linewidth]{layered_density.png}
        \caption{}
        \label{fig:offset_density}
  \end{subfigure}
  \begin{subfigure}{0.55\textwidth}
        \includegraphics[width=1\linewidth]{disordered_density.png}
        \caption{}
        \label{fig:offset_density}
  \end{subfigure}
  \caption{(a) Parallel Displaced (b) Sandwiched (c) "Disordered"}~\label{fig:densities}
  \end{figure}

>>>>

  While RO and NF have seen many advances in the past few decades, they 
  are far from perfect separation technologies.
  \begin{itemize}
    \item \textit{RO membranes}
    \item Inconsistent performance : Current state-of-the-art RO membranes are unstructured with
    tortuous and polydisperse diffusion pathways which leads to 
    inconsistent performance \cite{song_nano_2011}
    \item High energy requirements : Necessarily high feed pressures 
    drive up energy requirements which strains developing regions and
    contributes strongly to CO\textsubscript{2} emissions. \cite{mcginnis_global_2008}
    \item Separation based on differences in solubility and diffusivity:
%MRS2: couldn't you argue that the fact that they diffuse and dissolve at different rates means that it makes it possible to design targeted separations?
%BJC: right, I say something like that in the sentence following this one. It would be more challenging to design a targeted separation 
% especially if the feed is complex
    Moreover, designing RO membranes to achieve targeted separations of 
    specific solutes in a complex feed solution is nearly impossible because various solutes dissolve
    into and diffuse through the polymer matrix at different rates. \cite{wijmans_solution-diffusion_1995}
    \item With optimization, one can exploit these differences to create a functional
    selective barrier.
    \item \textit{NF membranes}
    \item NF was introduced as an intermediate between RO and ultrafiltration,
    having the ability to separate organic matter and salts on the order of 
    one nanometer in size.
%MRS2: is ``well-defined'' the right word if there is a ``pore size distribution''?
%BJC: By well-defined, I'm thinking of cylinders versus polymer matrix. One is definitely more
% defined than the other. What is the right word for that? unambiguous? explicit?
%MRS3: hmm.  Maybe look and see what other people call them when describing NF membranes.
    \item Larger and well-defined pores drive down energy requirements while
    still affording separation of solutes as small as ions to some degree \cite{van_der_bruggen_review_2003}
    \item NF is often used as a precursor to reverse osmosis
    \item Unfortunately, NF membranes, like RO, possess a pore size 
    distribution which limits their ability to perform precise separations \cite{bowen_modelling_2002}
  \end{itemize}

@@

  A molecular level understanding of LLC membrane structure, enabled by molecular
  dynamics simulations, will provide guidelines to reduce the large chemical space
  available to design monomers for creation of separation-specific membranes.
  \begin{itemize}
    \item Over the past 20 years, H\textsubscript{II} phase LLC membrane studies have been limited primarily
    to Na-GA3C11 with some characterization done after minor structural modifications
    \item Resel et al. varied the length of the monomer tails and the counterion used and 
    observed its affect on pore spacing \cite{resel_structural_2000}.
    \item Rejection studies show that this membrane can not perform separations of
    solutes less than 1.2 nm in diameter because the pores are too large \cite{zhou_supported_2005}.
    \item  We do not yet understand how to controllably reduce the effective pore size or 
    how to tune the chemical environment in the nanopores for effective water
    desalination or small organic molecule separations.
    \item It will be challenging to efficiently narrow down the large design space in 
    a laboratory setting without a robust model.
    \item The only source of predictive modeling for LLC systems have been macroscopic models 
    which likely do not adequately describe transport at these length scales. % BJC: Reference. I think it is just w.r.t. bicontinuous cubic
%MRS2: not quite sure of the phrasing here on next two points. Something more a long the lines of the choice of head group 
%may allow tuning rejection of solutes based on size?
%BJC: okay tried again
    \item The choice of head group may allow us to tune pore size for size exclusion driven separations
%MRS3: choice of counterion also affects size.  Counterion type probably doesn't affect Donnan potential that much IRCC, more the total amount of counterion.
    \item Choice of counterion may influence the establishment of a Donnan potential
    affecting the degree to which the membrane can exclude charged species.
    \item A good molecular model should incorporate a detailed picture of the nanoscopic pore 
    structure which will be crucial to understanding the role of monomer structure in 
    membrane design.
    \item Molecular dynamics simulations will provide the required level of detail
  \end{itemize}

  %BJC: Figure here showing previous level of detail compared to atomistic level of detail
  %MRS2: provenance of the atomistic figure could be explained better in the caption. It also looks disordered. Will eventually need a higher resolution figure for the previous understanding pore.
  %BJC: Okay. The picture here is of the "ordered pore" phase. It looks disorderd from this angle. I could use a different slice of the membrane to see if it looks any better
%MRS: if this is ordered, OK.  Maybe could try to rotate it a bit more towards the reader.
  \begin{figure}
  \centering
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[width=\textwidth]{nanopore_undetailed.jpg}
		\caption{}~\label{fig:undetailed_pore}
	\end{subfigure}
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[width=\textwidth]{detailed_pore.png}
		\caption{}~\label{fig:detailed_pore}
	\end{subfigure}
  \caption{(a) Our previous understanding of the pore structure allows us to speculate
	   about separation behavior. (b) A detailed molecular model will allow us to
	   directly observe solute transport. Here, four stacked layers of 5 monomers
           are pictured atomistically. The hydrophilic region is in red and the 
	   hydrophoic region is colored blue.}~\label{fig:detail}
  \end{figure}

@@

  Despite having structural data, there is still information which
  experiment cannot definitively answer. There are several key
  questions that we intend to answer 
%MRS3: we can't ``answer'' for sure, since it's a model, and may omit things. Could put ``several key questions we will investigate'' or ``hope to provide evidence for or against'', or something like that.
which will be laid out and numbered
  in subsequent paragraphs.

@@

%MRS: break this down into more hierarchy - there are 11 separate points.  Can they be grouped at all for clearer understanding? elucidate what the aspects are we need a clearer picture. 
%BJC: I moved the logic for choice of monomer to system to model to a separate paragraph (above this). I then moved the other three things I talk about into their own paragraphs to keep the ideas separate (following paragraphs).
%MRS: since the questions we answer in the paper have a lot to do with the structure of the membranes, somewhere relatively early in the paper we need to lay out the facts that 1) these membranes could be really good (I think you do establish this) 2) we really would need to know the structure better to really do useful things with these, and 3) MD is a very helpful way to gain more information about the structure, because of the limited range of experiments we can do with these systems. That is the main rationale, and it should be made clear fairly early in the introduction.  Very important to make your story absolutely as clear as possible.
%BJC: I moved and edited a paragraph that was originally later on to be 2 above this one. I think it does a better job addressing points 2 and 3. And now it addresses those points earlier in the intro (as early as it could go where I think it still makes sense in context)

@@

  Monomers in the Col\textsubscript{h} system are theorized to be partitioned
  into stacked layers which form columnar pores. There has been no definitive
  answer in literature regarding the number of monomers in each layer. 
%MRS3: the intro to the list talks about only ``numbers of monomers in each layer'', whereas the list below talks about several questions relating to layers. Can you rephrase so it's more general? Maybe just take out that sentence above.
We want to
  know 
\begin{enumerate}
\item If layers do exist, how many monomers constitute a single layer? \label{point:monomernum}
  \begin{itemize}
    \item A simple molecular simulation study of a similar molecule suggested
    that there are 4 monomers in each layer. Their estimation is based on a
    simulated system containing only 16 total monomers which likely does not sufficiently
    model the chemical environment present in the real system.~\cite{zhu_methacrylated_2006}. 
    \item A separate calculation based on the volume of the liquid crystal monomers proposes
    that there are seven monomers in each layer~\cite{resel_structural_2000}. 
    \item A molecular model orders of magnitude larger than any other reported atomistic 
    liquid crystal membrane simulations has the best chance of directly answering this question.
    \item We can directly change the layer composition and note its effect on membrane structure.
  \end{itemize}

@@

  \item Is it necessary to include any water in order to appropriately model the 
  Col\textsubscript{h} phase? \label{point:water}
  \begin{itemize}
	\item While the Col\textsubscript{h} phase is described as dry, it is 
          %MRS3: ``possible'' -> likely, and give some explanations why we think likely?
	possible that small amounts of ambient water may be leached into the system.
	\item The hydrogen bonding network formed by the water may play a role 
        in structuring the pore.
	\item We can use simulated X-ray diffraction patterns to see if there is
        any meaningful structural difference between a "dry" and "wet" system
  \end{itemize}
  \end{enumerate}

@@

  We chose the layer spacing for the initial configuration based on R-$\pi$ 
  and then allowed system to readjust during equilibration.
  \begin{itemize}  
    \item Each monomer was rotated so the plane of the aromatic head groups would
    be coplanar with the xy plane.
    \item We explore two different initial layer spacings
    \item The first is exactly equal to R-$\pi$ with layers placed so aromatic
    rings are stacked 3.7 \angstrom~ apart in the z-direction.
    \item A second system is explored with an initial layer spacing of 5 \angstrom
%MRS3: at least state in the main body what went wrong with it in a sentence or so, with details in supplemental. 
   \item A third system with an initial layer spacing of 10 \angstrom was briefly
    explored and the reader can learn more about it in the Supplemental Information.
% BJC: Move all this lower down to be with the rest of the equilibrium calculations
%    \item To quantify the degree of layering in our system, we calculated
%    a spatial correlation function, $g(z)$, measured along the z-axis 
%    (perpendicular to the membrane plane)
%    \item To calculate $g(z)$, we binned the z component distances between 
%    the center of mass of each benzene ring and all others of the same pore 
%    over 50 ns of equilibrated trajectory and then normalized by the average
%    number density.
%    \item We compare the degree of layering between systems based on the
%    amplitude of the first peak in $g(z)$. We halve the difference between the
%    maximum of the first peak and the following minimum. We compare the difference
%    to the mean to get a percentage deviation from the average number density.
%    \item To extract the average distance between layers we applied a discrete 
%    fourier transform to $g(z)$ and extracted the highest intensity frequency 
  \end{itemize}

@@

  We developed equilibration schemes to create dry and wet configurations. Both schemes 
  start with an initial configuration generated according to the previous guidelines.
  \begin{itemize}
      \item Dry equilibration scheme:
      \begin{itemize}
        %MRS2: what initial force constant, and does it matter (that can be in supplemental) 
	%BJC: I've never explored whether there is any dependence on initial force constant. It probably doesn't matter unless it starts really low
        %MRS3: then just state what you did, so it could be repeated. 
          \item Restraints fix monomer head groups in the sandwiched or parallel-displaced
          configurations while allowing monomer tails to settle.
          \item Doing so also mitigates system dependence on initial monomer configuration.
          \item The restrained portion of the equilibration scheme is run in the NVT ensemble.
          \item Every 50 ps, we reduce the force constants by the square root of its
          previous value, starting from 1e6 KJ mol$^-1$ nm$^-2$.
          \item Once the force constant is below 10 KJ mol$^-1$ nm$^-2$, the restraints are
          released linearly until there is no more restraining potential.
          \item The resulting unrestrained structure is allowed to equilibrate for 5 ns 
	  in the NPT ensemble with pressure controlled by the berendsen barostat.
	  \item Long, NPT equilibration simulations are run for at least 400 ns using the 
          Parrinello-Rahman barostat with a time consant of 10 ps.
      \end{itemize}
      \item Wet equilibration scheme: 
      \begin{itemize}
	  % BJC: the following three bullet points ended up not being the case since I used significantly less water
          %\item We obtained an estimate of equilibrium pore water content by solvating
          % an initial configuration with water baths above and below the membrane. 
	  %\item We ran the solvated system according to Scheme A but with a 1000 ns equilibration
          %\item Using that number we added water to the pores of an initial 
	  %configuration and equilibrated according to scheme B. 
	  \item In order to create a wet system, we solvated an initial configuration with water using gmx solvate
          \item All water molecules placed outside the pore region are removed 
	  \item Waters inside the pore region are randomly removed until the desired concentration
	  of water in the pores is reached. 
	  \item The remainder of the equilibration is the same as the dry system
      \end{itemize}
      \item In all cases, the v-rescale thermostat was used with tau-t = 0.1 % BJC: supplemental
  \end{itemize}

@@

  To calculate the equilibrated pore spacing, we measured the distance between pore
  centers.
  \begin{itemize}
	\item Pore centers were located by averaging the coordinates of sodium ions in their
	respective pores.
	\item Statistics were generated using the bootstrapping technique (See Supplemental Information)
	% Supplemental vvvv
        \begin{itemize}
        	%MRS2: make clear sampling what?
	        \item We are interested in 5 pore-to-pore distances which should all be equal in
	        a perfect hexagonal array, however only 4 distances are independent % can visualize this in supplemental material. Will make a lot of things more clear below.
	        \item Each pore spacing has its own trajectory of spacing vs. time. Using data collected after
	        the system is equilibrated, we calculate how long it takes for the data in each of the 5
	        trajetories to become uncorrelated using pymbar.timeseries.integratedAutocorrelationTime() % reference
	        \item We break the full trajectories down into sub-trajectories based on the
	        maximum autocorrelation time of those found in the previous step.
	        \item For each bootstrap trial, we recreate an equilibrium trajectory by randomly
	        sampling pore spacings from the sub-trajectories
	        \item We get an average value for each pore spacing by finding the mean of the bootstrapped data
	        \item We calculate the overall average as the mean of all bootstrapped pore spacings
                  %MRS3: this sounds a little convoluted below.  Should just be the standard error in the boostrap estimates of the pore spacing?
                  %MRS3: this can be fixed in the draft not outline.
	        \item The uncertainty for each pore spacing is calculated as $\dfrac{<x> - x}{4}$ where $<x>$ is the
	        average spacing from the bootstrap trial and $x$ is the average value of one of the pore spacings.
	        \item We report the mean of these uncertainties
	        % MRS: need to deal with the fact only 4 are independent in the bootstrapping?  How did we do that?
	        % BJC: Updated
        \end{itemize}
        % Supplemental ^^^^^
  \end{itemize}

@@

  \begin{figure}
  \centering
  	\begin{subfigure}{\linewidth}
	\centering
		\begin{subfigure}{0.45\textwidth}
        		\centering
        		\includegraphics[width=\linewidth]{offset_tail_packing.png}
        		\caption{}~\label{fig:offset_tails}
		\end{subfigure}
		\begin{subfigure}{0.45\textwidth}
		\centering
	        	\includegraphics[width=\linewidth]{offset_angle_v_I.png}
		        \caption{}~\label{fig:offset_integration}
		\end{subfigure}
	\end{subfigure}
	\begin{subfigure}{\linewidth}
	\centering
		\begin{subfigure}{0.45\textwidth}
	        \centering
		        \includegraphics[width=\linewidth]{angles_traj_layered.png}
		        \caption{}~\label{fig:rz_layered}
		\end{subfigure}
		\begin{subfigure}{0.45\textwidth}
        	\centering
		        \includegraphics[width=\linewidth]{layered_angle_v_I.png}
		        \caption{}~\label{fig:layered_integration}
		\end{subfigure}
	\end{subfigure}
%MRS: 
  \caption{The distribution of angles w.r.t. the xy plane between alkane chain tail centroids and nearest
  neighbor centroids for equilibrated parallel displaced (a) and sandwiched (c) configurations. The
  same peaks are visible when the 2D simulated diffraction data is radially integrated in the R-alkanes region,
  (b) and (d) respectively.}~\label{fig:tail_packing}
%MRS2: not clear what the difference is between right and left side figures.
%BJC: That's good! One is angles measured between tail centroids and the other is radial integration of the simulated xrd patterns. I'll try make that more clear in the caption.
%MRS3: sorry, I meant to say not clear what the SOURCE of the data is in right vs. left.  Make sure colors and labels are consistent in final 
%MRS: not clear why some are angles with x axis, and some are angles w neighbors?  Shouldn't it be the same thing plotted in both graphs?
  \end{figure}

  The $z$-direction correlation functions show that layers in our model prefer
  to stack further apart than the 3.7 \angstrom~ suggested by experiment. We attempted equilibration
  with layers stacked 5.0 nm apart and to our surprise, we observed
  long-term stability of a qualitatively different configuration suggesting
  that we have found more than one metastable free energy basin, corroborating \ref{point:metastable}.
  %that the initial distance between layers can influence the approach towards 
  %equilibrium.
  \begin{itemize}
        \item Equilibrated systems built according to the 3.7 \angstrom~layer spacing
        implied by R-$\pi$ are characterized by a defined, cylindrical and open pore
        structure.
        \item We will refer to this large set of configurations, with an open pore, as
        the ordered pore basin open pore (Figure~\ref{fig:OPbasin}).
        \item Simulations of systems built with layers stacked 5
        \angstrom~ apart results in a pore structure characterized by high radial
        disorder, while still maintaining partitioning between hydrophobic and
        hydrophilic regions.
 %MRS3: need to address the similar RDFs.  What data can show they are different? Or maybe they HAVE rdfs, but not good layering? Have you done the fourier transform on them? We can still rule them out based on the XRD, of course.  But surprising the same RDFs.
        \item This will be called the disordered pore basin (Figure~\ref{fig:DPbasin}).
        \item This LLC membrane may exist in at least two metastable states.
        \item The distinct difference in pore structure exhibited by each phase will
        likely lead to different transport mechanisms.
        \item Because the pore structure will vary between each basin, understanding
        which phase exists experimentally is necessary in order to ensure we are
        studying the system which actually dominates.
  \end{itemize}

  \begin{figure}[!ht]
        \centering
        \begin{subfigure}[b]{0.475\textwidth}
                \centering
                \includegraphics[width=\textwidth]{280K_tramp_close.png}
                \caption{}\label{fig:OPbasin}
        \end{subfigure}
        \begin{subfigure}[b]{0.475\textwidth}
                \centering
                \includegraphics[width=\textwidth]{340K_tramp_close_full.png}
                \caption{}\label{fig:DPbasin}
        \end{subfigure}
        \caption{From a qualitative standpoint, the ordered pore basin (a) is characterized by a
        hollow cylindrical pore while pores in the disorder pore basin (b) are characterized by a
        higher degree of radial disorder}\label{fig:basins}
%MRS3: see note on RDF.
  \end{figure}

@@

  % BJC: I'm not sure what to make of this figure right now. The parallel-displaced configuration shares
  % more with the disordered pore phase
  % MRS3: yeah, see my notes on this.  wE need to understand how it can be visually open, and still have this structure.  
  % doesn't this structur have a different pore-to-pore distance.
  \begin{figure}
  \centering
  \begin{subfigure}{0.55\textwidth}
        \includegraphics[width=1\linewidth]{offset_density.png}
        \caption{}
        \label{fig:offset_density}
  \end{subfigure}
  \begin{subfigure}{0.55\textwidth}
        \includegraphics[width=1\linewidth]{layered_density.png}
        \caption{}
        \label{fig:offset_density}
  \end{subfigure}
  \begin{subfigure}{0.55\textwidth}
        \includegraphics[width=1\linewidth]{disordered_density.png}
        \caption{}
        \label{fig:offset_density}
  \end{subfigure}
  \caption{(a) Parallel Displaced (b) Sandwiched (c) ``Disordered''}~\label{fig:densities}
  \end{figure}

%MRS3: this needs a caption, and need to write some text analyzing it.  The thing that throws me off is that the RDF's look about the same for parallel displaced and disordered!  That's pretty hard to rationalize with the idea that the pores are open vs. closed. What am I missing? Very interesting regardless!
